MCC Novelty Kids

### 2018-10-29 stats/mcc_inf_pilot_kids.Rmd
adding kids exp 3

---
title: "MCC Informativeness - moving agent pilots kids"
author: "Manuel Bohn"
date: "7 2 2018"
output: html_document
---

```{r setup}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)

## importing data
d <- inf.data %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

Sanity checks.
```{r training performance}
## performance in training and test
d %>%
  group_by(trial_type) %>%
  summarise(n = length(unique(subid)), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 3)

# Plot with 95%CI
```{r plot}

p <- d %>%
  filter(trial_type == "test") %>%
  group_by(control, subid) %>%
  summarise(correct = mean(correct))

@@

Effect of target object location on table. Inner = closer to the agent, outer = further away. 
For test this does not seem to matter. For control: If they had a tendency to pick the one closer to the agent, they should be above chance for inner and below chance for outer. Only the latter seems to be the case. Not sure why, could also be just noise. 
```{r inner vs outer location}
## performance in training and test
ms <- d %>%
  filter(trial_type == "test") %>%
  group_by(targetOnTable, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = targetOnTable, y = mean, fill = targetOnTable)) +
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few(base_size = 12) +
  ylim(0,1)+
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_fill_solarized(name="Trial Type",
                     breaks=c("false", "true"),
                     labels=c("Test", "Control"))

@@

Conditions by trial.
```{r plot trials}
ms <- d %>%
  filter(trial_type == "test") %>%
  group_by(control, trial, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = trial, y = mean, col = control)) +
  geom_line(aes(group= control)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  theme_few() + 
  scale_colour_solarized()
```

Conditions by agents.
```{r trials per condition and agent}
d %>%
  filter(trial_type == "test") %>%
  group_by(agent,control) %>%
  summarise(n = length(agent), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)
```

Agent effect. Dog seems a little strange (despite enough data) but no idea what to make of this.
```{r plot agents}
ms <- d %>%
  filter(trial_type == "test") %>%
  group_by(control, agent, id) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = agent, y = mean, col = control)) +
  geom_line(aes(group= control)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  theme_few() + 
  scale_colour_solarized()
```

T-tests against chance test and control
```{r}
## t tests against chance
d %>%
  filter(trial_type == "test") %>%
  group_by(control, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(control) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(control,df,t_value,p_value) %>%
  knitr::kable(digits = 3)

Agent can be seen as an approximation for "item".

```{r}
## model
library(lme4)

data_inf <- d%>%
  filter(trial_type == "test")%>%
  mutate(trial = scale(as.numeric(trial), center = TRUE, scale=FALSE))

# registered  model
lm <- glmer(correct ~ 1 +
              (1 |id) + (1 |agent), 
              data = data_inf, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(lm)
```

>>>>

---
title: "MCC Informativeness - Pilot with Kids"
author: "Manuel Bohn"
output: html_document
---

```{r setup, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)

## importing data
files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/trees")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/trees/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_trees <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2",
         subage != "Test")

files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_4_map")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/robots_point_4_map/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_robots <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2",
         subage != "Test")

files <- dir("~/Work/MCC/git-mcc/kids_info_pilot/trees_left_right")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_pilot/trees_left_right/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

d_trees_left_right <- raw_data %>%
  filter(trial != "train1",
         trial != "train2",
         trial != "filler1",
         trial != "filler2",
         subage != "Test")
```

# Sample

## Trees

```{r sample trees, echo = F}

d_trees %>%
  group_by(subage) %>%
  summarise(n = length(unique(subid))) %>%
  knitr::kable(digits = 3)
```

## Robots

```{r sample robots, echo = F}

d_robots %>%
  group_by(subage) %>%
  summarise(n = length(unique(subid))) %>%
  knitr::kable(digits = 3)
```

## Trees left right

```{r sample trees left right, echo = F}
d_trees_left_right %>%
  group_by(subage) %>%
  summarise(n = length(unique(subid))) %>%
  knitr::kable(digits = 3)
```

# Results overall

## Trees

Experiment: [Tree Setup](https://langcog.stanford.edu/expts/cue/exp/kids/kids_cue_barrier7_tree.html)

```{r , echo = F}

p <- d_trees %>%
  group_by(control,subid) %>%
  summarise(correct = mean(correct))

@@

## Robots

Experiment: [Robot Setup](https://langcog.stanford.edu/expts/cue/exp/kids/kids_cue_barrier6.html)

```{r, echo = F}

p <- d_robots %>%
  group_by(control,subid) %>%
  summarise(correct = mean(correct))

## Trees left right

Experiment: [Trees with objects left and right](https://langcog.stanford.edu/expts/cue/exp/kids/kids_cue_barrier8_tree_left_right.html)

```{r, echo = F}

p <- d_trees_left_right %>%
  group_by(control,subid) %>%
  summarise(correct = mean(correct))

# Results by age

## Trees

```{r , echo = F}

p <- d_trees %>%
  group_by( subage, control, subid) %>%
  summarise(correct = mean(correct))

ggplot() +
  geom_jitter(data = p, aes(x = control, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = control, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

## Robots

```{r , echo = F}

p <- d_robots %>%
  group_by( subage, control, subid) %>%
  summarise(correct = mean(correct))

ggplot() +
  geom_jitter(data = p, aes(x = control, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = control, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

## Trees left right

```{r , echo = F}

p <- d_trees_left_right %>%
  group_by( subage, control, subid) %>%
  summarise(correct = mean(correct))

ggplot() +
  geom_jitter(data = p, aes(x = control, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = control, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Positioning bias (top vs. bottom object more informative)

## Trees
```{r , echo = F}

p <- d_trees %>%
  group_by( subage, targetPosition, subid) %>%
  summarise(correct = mean(correct))

ggplot() +
  geom_jitter(data = p, aes(x = targetPosition, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = targetPosition, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

## Robots

```{r , echo = F}

p <- d_robots %>%
  group_by( subage, targetPosition, subid) %>%
  summarise(correct = mean(correct))

ggplot() +
  geom_jitter(data = p, aes(x = targetPosition, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = targetPosition, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

## Trees left right (here it is inner vs outer object)

```{r , echo = F}

p <- d_trees_left_right %>%
  group_by( subage, targetPosition, subid) %>%
  summarise(correct = mean(correct))

ggplot() +
  geom_jitter(data = p, aes(x = targetPosition, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p1, aes(x = targetPosition, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

```{r}
# power analysis for cue barrier

xx <- bind_rows(d_trees,
d_trees_left_right,
d_robots)%>%
  filter(subage != "2")

 power_data <- xx %>%
  group_by(subage,subid) %>%
  summarise(correct = mean(correct))

library("lsr")

library("pwr")

d_robots %>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(subage) %>%
  mutate(mean = mean(unlist(correct)),
        df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value,
        d = cohensD(unlist(correct), mu = 0.5)) %>%
  select(subage,mean,df,t_value,p_value,d) %>%
  knitr::kable(digits = 3)

cohensD(power_data$correct,mu = 0.5)

pwr.t.test(d = 0.65, power = 0.8, type = "one.sample")

pwr.t.test(d = 0.8, power = 0.8, type = "one.sample")

pwr.t.test(d = 1.05, power = 0.8, type = "one.sample")

@@

### 2018-10-29 stats/mcc_novel.kids.Rmd
adding kids exp 3

d<- read_csv(file="kids_novel.data.csv")%>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

# check if someone needs to be excluded because wrong in training
# no one
d %>%
  filter(trial_type == "train") %>%
  group_by(subid)%>%
  summarise(correct = mean(correct)) %>%
  filter(correct == 0)

@@

## sanity checks
d %>%
  group_by(trial_type) %>%
  summarise(n = length(unique(subid)), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)

@@

d %>%
  group_by(age)%>%
  summarise(n = length(unique(subid)),
            age_num = mean(age_num))%>%
  knitr::kable(digits = 2)
```

@@

```{r plot across age}
p3 <- d %>%
  filter(trial_type == "test") %>%
  group_by(change, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = p3, aes(x = change, y = correct, col = change, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p4, aes(x = change, y = mean, col = change, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

@@

The following two plots suggest some sort of linear developmental trajectory.

```{r plot by age}
p1 <- d %>%
  filter(trial_type == "test") %>%
  group_by(change,age, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = p1, aes(x = change, y = correct, col = change, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = change, y = mean, col = change, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~age)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

@@

```{r plot by numerical age}
p5 <- d %>%
  filter(trial_type == "test") %>%
  group_by(change,age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p5, aes(x = age_num, y = correct, col = change)) +
  geom_smooth(method = "loess", se = T)+
  geom_jitter( alpha = .5, height = 0.04)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

>>>>

d_new <- read_csv(file="kids_novel.data.csv")%>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

d_old <- read_csv(file="kids_novel.data.round1.csv")%>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

# check if someone needs to be excluded because wrong in training
# no one
d_new %>%
  filter(trial_type == "train") %>%
  group_by(subid)%>%
  summarise(correct = mean(correct)) %>%
  filter(correct == 0)

d_old %>%
  filter(trial_type == "train") %>%
  group_by(subid)%>%
  summarise(correct = mean(correct)) %>%
  filter(correct == 0)

d <- bind_rows(d_new,d_old)

@@

## sanity checks
d %>%
  group_by(condition,trial_type) %>%
  summarise(n = length(unique(subid)), 
            correct = mean(correct)) %>%
  knitr::kable(digits = 2)

@@

d %>%
  group_by(condition,age)%>%
  summarise(n = length(unique(subid)),
            age_num = mean(age_num))%>%
  knitr::kable(digits = 2)
```

@@

```{r plot across age}
p3 <- d %>%
  filter(trial_type == "test") %>%
  group_by(condition,change, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = p3, aes(x = change, y = correct, col = change, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p4, aes(x = change, y = mean, col = change, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  facet_grid(condition ~.)+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

@@

```{r plot by age}
p1 <- d %>%
  filter(trial_type == "test") %>%
  group_by(condition,change,age, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = p1, aes(x = change, y = correct, col = change, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = change, y = mean, col = change, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(condition~age)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

@@

```{r plot by numerical age}
p5 <- d %>%
  filter(trial_type == "test") %>%
  group_by(condition,change,age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p5, aes(x = age_num, y = correct, col = change)) +
  geom_smooth(method = "loess", se = T)+
  geom_jitter( alpha = .5, height = 0.04)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
   facet_grid(condition ~.)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```
